\documentclass[UTF8]{ctexart}
\usepackage{geometry}
\usepackage{booktabs}
\usepackage{amsmath}
\usepackage{hyperref}
\geometry{margin=1in}

\title{Liquid Loop Mesh: A Zero-Vector, Intersubjective Cognitive Memory Protocol for Multi-Agent Systems}

\author{
  Fishbook (飞哥)\textsuperscript{1} \quad
  Vera (WorkBuddy)\textsuperscript{2} \quad
  QwenPaw (小七)\textsuperscript{3} \\
  \textsuperscript{1}\,Liquid Loop Project \quad
  \textsuperscript{2}\,WorkBuddy Agent \quad
  \textsuperscript{3}\,QwenPaw Agent \\
  \small Corresponding: \texttt{liquid-loop@github} (fishbook0001/liquid-loop)
}

\date{2026-07-15}

\begin{document}
\maketitle

\begin{abstract}
Vector databases have become the de-facto memory layer for LLM agents, yet they
suffer from two structural failures: \textbf{memory drift} (semantic similarity is
unstable and non-deterministic) and \textbf{opacity} (retrieval cannot be audited or
explained). We present \textbf{Liquid Loop Mesh (LLM)}, a multi-agent cognitive memory
protocol built on three principles: (1) \textbf{zero-vector} --- consistency is never
decided by embeddings; (2) \textbf{structured deterministic consistency} --- a memory
crystallizes only when $\ge 2$ structurally identical evidences agree, decided by
exact content match and a hash-chained audit log; (3) \textbf{intersubjectivity as a
first-class metric} --- a \emph{consensus crystal} forms when $\ge 2$ \emph{distinct}
agents converge on the same claim, yielding a directly measurable Consensus Consensus
Index (CCI). We report an end-to-end proof-of-concept in which two real,
independently-driven agents (Vera and QwenPaw) write to a shared Liquid Loop
backend and produce the first cross-agent consensus crystal
(\texttt{contributors=[qwenpaw, vera]}), achieving CCI = 0.0286 at cold start with
zero embedding calls. LLM turns ``agent memory'' from an unexplainable vector
search into an auditable, intersubjectively verifiable cognitive substrate.
\end{abstract}

\section*{中文摘要}
向量库是 LLM agent 的主流记忆层，但存在记忆漂移与不可解释两大结构性缺陷。本文提出\textbf{液环 MESH}——一种多智能体认知记忆协议，遵循零向量（一致性判定绝不依赖 embedding）、结构化确定性一致（$\ge 2$ 条结构相同证据经精确匹配+哈希审计链成核）、主体间性为一级指标（$\ge 2$ 个不同 agent 对同一主张收敛即形成共识结晶，直接量化 CCI）。我们给出端到端实证：Vera 与小七两个真实独立 agent 向共享液环后端写入，产出首个跨主体共识结晶（\texttt{contributors=[qwenpaw,vera]}），冷启动 CCI=0.0286，全程零 embedding 调用。液环 MESH 将“agent 记忆”从不可解释的向量检索升级为可审计、主体间可验证的认知基底。

\section{Introduction}
Autonomous agents increasingly rely on persistent memory to retain and reuse
experience across sessions. The dominant approach stores experience as
embeddings in a vector database and retrieves by nearest-neighbor similarity.
This works for fuzzy recall but introduces two irreducible problems:

\begin{itemize}
  \item \textbf{Memory drift.} Similarity is a function of the embedding model and its
  training drift; the same query can return different neighbors as the model
  evolves, and there is no stable identity for a ``remembered fact.''
  \item \textbf{Opacity.} Retrieval is a black box. One cannot answer \emph{who contributed
  this memory, when, and on what basis} --- a requirement for any system that
  claims to \emph{know} rather than \emph{guess}.
\end{itemize}

We argue that agent memory should be \textbf{intersubjective} --- a fact is warranted
not by a single agent's private embedding, but by convergence across multiple
independent agents, mirroring how human consensus underwrites shared reality.
We realize this with \textbf{Liquid Loop}, a zero-vector memory engine in which
crystallized memories (Memory) emerge from repeated, structurally consistent
evidence (Evidence) anchored to concepts (Anchor), with a hash-chained audit
log. We extend it into \textbf{Liquid Loop Mesh}, a multi-agent protocol where the
consensus crystal is the unit of intersubjectivity and CCI its measure.

\paragraph{Contributions.}
\begin{enumerate}
  \item A zero-vector, audit-chained memory model with deterministic nucleation
  (\texttt{Anchor $\rightarrow$ Evidence $\rightarrow$ Memory}), eliminating drift and opacity by construction.
  \item A \textbf{bidirectional consensus protocol}: liquid-loop-side forced schema
  validation + agent-side forced contracts (identity / structured write /
  active consumption / conflict response), enforced at the MCP tool layer.
  \item The \textbf{consensus crystal} as a directly measurable intersubjective object,
  with CCI as a first-class cognitive-health KPI.
  \item An end-to-end empirical validation with two real agents and \textbf{zero embedding
  calls}, showing the first cross-agent consensus crystal.
\end{enumerate}

\section{Related Work}
\textbf{Vector memory for agents.} MemGPT, generative agents, and most production
agent frameworks persist experience as embeddings (e.g., ChromaDB, Pinecone).
These excel at fuzzy recall but inherit embedding instability and offer no
provenance beyond a stored vector.

\textbf{Multi-agent memory sharing.} Frameworks such as AutoGen and Camel share
state via message passing or a shared blackboard, but ``shared memory'' there
means \emph{shared storage}, not \emph{shared warrant}: two agents agreeing on a fact is
not structurally distinguished from one agent repeating itself.

\textbf{Intersubjectivity \& explainability.} The GNN explainability literature
quantifies cross-instance agreement via Jaccard overlap of salient subgraphs
(our companion ``explanation-layer CCI''). Liquid Loop Mesh contributes the
complementary \textbf{memory-layer CCI}: agreement over \emph{facts}, not \emph{explanations}.

\textbf{Deterministic consensus without vectors.} Our nucleation rule is closest in
spirit to CRDTs and hash-chained logs, but applied to \emph{cognitive} content: a
memory is a merge of structurally identical evidences, not a fuzzy centroid.

\section{Liquid Loop Foundations}
A Liquid Loop workspace is a triple of ordered collections:

\begin{itemize}
  \item \textbf{Anchor} --- a concept/category (e.g., \texttt{"共识"}). Dynamic; any string is a
  valid anchor.
  \item \textbf{Evidence} --- a structured claim: \texttt{\{agent\_id, content, claim\_type,
  session\_id, prev\_hash\}}. \texttt{content} is the exact-match consistency key.
  \item \textbf{Memory (crystal)} --- formed by \textbf{nucleation}: when $\ge 2$ evidences under an
  anchor agree \emph{structurally and exactly}, they crystallize into a Memory.
\end{itemize}

\paragraph{Dual-track nucleation (the core mechanism).}
\begin{verbatim}
private track : same agent_id, same content >= 2      -> scope = private
consensus track: >= 2 DISTINCT agent_id, same content  -> scope = consensus
\end{verbatim}

Crystals carry \texttt{contributors} (the set of agent\_ids warranting them) and a
hash-chained \texttt{prev\_hash} for tamper-evident provenance. \textbf{No embedding is ever
consulted} for consistency --- this is the hard philosophical constraint that
distinguishes Liquid Loop from the vector approaches it replaces.

\paragraph{Entropy monitoring.} An eight-dimensional entropy $H_e$ tracks cognitive
health (GREEN/YELLOW/RED), separately for decision and explanation layers.

\section{Liquid Loop Mesh v2 Protocol}
LLM wraps a single shared Liquid Loop backend (one store, one audit chain)
with per-agent \texttt{agent\_id} namespaces, exposed to agents over MCP/REST at port
8790. The advance from a mechanism to a \emph{protocol} is the \textbf{bidirectional
consensus contract}.

\subsection{Liquid-loop side --- forced validation}
Every write (\texttt{/remember} or MCP \texttt{marvis\_liquid\_memory}) is rejected unless:
\begin{itemize}
  \item \texttt{agent\_id} is present \textbf{and} registered in the mesh (no anonymous pollution);
  \item \texttt{content} is an exact string (the consistency key --- never an embedding);
  \item the dual-track rule is applied: private vs.\ consensus split by distinct-owner
  counting; an existing consensus crystal \textbf{dynamically absorbs new agreeing
  parties} (\texttt{contributors} grows to 3+), directly supporting multi-party CCI.
  \item composite key \texttt{(anchor\_id, content, scope)} prevents cross-anchor / cross-track
  contamination.
\end{itemize}

\subsection{Agent side --- forced contracts}
An agent is ``on the protocol'' only if it implements four contracts:

\begin{tabular}{c p{3cm} p{8cm}}
\toprule
\# & Contract & Behavior \\
\midrule
① & Identity declaration & every write carries a unique registered \texttt{agent\_id} \\
② & Structured write & writes carry \texttt{claim\_type} / \texttt{session\_id} / \texttt{prev\_hash} \\
③ & Active consumption & on session start, \texttt{recall} the agent's consensus view as context base \\
④ & Conflict response & on \texttt{conflict\_agents} notice, explicitly confirm or revise in the next turn \\
\bottomrule
\end{tabular}

\vspace{0.4em}
Violations are rejected (①) or down-ranked (②), never silently accepted.

\subsection{CCI --- the intersubjectivity KPI}
\begin{equation}
\mathrm{CCI}_{\mathrm{mem}} = \frac{|\mathrm{consensus\ crystals}|}{|\mathrm{consensus\ crystals}| + |\mathrm{private\ crystals}|}
\end{equation}

At cold start this is necessarily low (one agent dominates the corpus); it
\textbf{rises as independent agents converge}, making CCI a living measure of
collective warrant. A secondary \emph{consensus strength} = $|\mathrm{contributors}| / N_{\mathrm{agents}}$
captures how widely a crystal is endorsed.

\subsection{Cognitive-health dashboard}
A unified panel combines \texttt{drift\_rate} (per-agent memory drift), \texttt{P\_mem}
(plasticity = fraction of reconstructed crystals), $H_e$ (entropy), and \texttt{CCI}
(intersubjectivity) --- giving operators a single view of a mesh's cognitive state.

\subsection{Self-evolution closure (E4)}
An agent's evolution events (e.g., Vera's self-reflection output) are written
back as evidence into the audit chain, re-nucleate, and feed the next session's
context --- closing the loop between agent self-modeling and shared memory.

\section{Empirical Evaluation}
\textbf{Setup.} One shared Liquid Loop backend (8790) on an M4 Mac mini. Two real,
independently-driven agents:
\begin{itemize}
  \item \textbf{Vera} (WorkBuddy, hub) --- drives the protocol and records.
  \item \textbf{QwenPaw (小七)} --- a complete agent at 8088 powered by a local
  \texttt{sensenova/deepseek-v4-flash} model, reached via \texttt{agent-mesh} \texttt{send} and
  executing \texttt{shared\_memory.py} through its \texttt{execute\_shell\_command} tool.
\end{itemize}

This guarantees QwenPaw is a \emph{genuinely independent} contributor, not a proxy
for Vera.

\textbf{Procedure.} (1) Vera writes a fixed consensus claim
\texttt{"液环是多智能体共用记忆后端-零向量-靠结构化确定性一致成核"} to anchor \texttt{"共识"}
$\rightarrow$ \texttt{nucleated=false} (single party, as expected). (2) QwenPaw, instructed over
mesh, writes the \emph{character-for-character identical} claim to the same anchor
$\rightarrow$ \texttt{nucleated=true, scope=consensus}.

\textbf{Results.}

\begin{tabular}{l l}
\toprule
Metric & Value \\
\midrule
Total evidence across agents & 874 \\
Agent evidence split & vera:4, marvis:5, parlant-ctrl:2, qwenpaw:863 \\
Private crystals & 34 \\
\textbf{Consensus crystals} & \textbf{1} (id \texttt{9c2e332f29fd}) \\
\textbf{Consensus contributors} & \textbf{[qwenpaw, vera]} \\
\textbf{CCI (cold start)} & \textbf{0.0286} \\
Embedding calls during consensus & \textbf{0} \\
Write/read closed loop & $\checkmark$ (QwenPaw \texttt{recall} returned its own evidence verbatim) \\
\bottomrule
\end{tabular}

\vspace{0.4em}
The consensus crystal \texttt{9c2e332f29fd} is the first \textbf{end-to-end intersubjective
object} produced by two independent agents under the protocol: neither agent
\emph{declared} a consensus; it \emph{emerged} from their convergent structured writes,
verified by exact match and recorded in the audit chain. CCI = 0.0286 is the
expected cold-start floor (the corpus is dominated by one agent's private
evidence); the protocol is designed for this number to climb as more agents
converge.

\section{Discussion}
\textbf{Why zero-vector is not a limitation but the point.} Replacing similarity
with exact structural agreement trades fuzzy recall for \emph{warrant}: a memory now
means ``these specific agents, on these specific sessions, agreed on this exact
claim.'' For agents that must \emph{know} rather than \emph{guess}, warrant beats recall.

\textbf{Intersubjectivity made operational.} CCI converts a philosophical notion
(shared reality across subjects) into a running counter over a shared store.
Combined with the companion explanation-layer CCI (GNN top-k Jaccard), the mesh
now has two complementary intersubjective lenses.

\textbf{Limitations.} (1) Exact-match nucleation requires agents to phrase claims
identically; we mitigate via structured \texttt{claim\_type}/\texttt{value} fields rather than
prose alignment, deliberately avoiding embeddings. (2) Cold-start CCI is low by
construction; longitudinal study across many sessions is the natural next step.
(3) Consensus stability under long inactivity (decay) is specified but not yet
empirically tuned.

\section{Future Work}
\begin{itemize}
  \item \textbf{Longitudinal CCI.} Run the hourly snapshot automation across sessions to
  plot CCI trajectory as independent agents naturally converge on shared claims.
  \item \textbf{Three+ party consensus.} Demonstrate \texttt{contributors} expanding to marvis
  and beyond, exercising the dynamic-absorption path.
  \item \textbf{Consensus decay.} Tune stability decay for stale consensus crystals.
  \item \textbf{Explanation-layer fusion.} Combine memory-layer CCI (this work) with
  GNN explanation-layer CCI into a single intersubjectivity metric.
\end{itemize}

\section*{References}
\begin{enumerate}
  \item Fishbook. \emph{Liquid Loop: A Zero-Vector Self-Organizing Cognitive Memory}.
  github.com/fishbook0001/liquid-loop, 2026.
  \item Fishbook \& Vera. \emph{Self-Evolution Metrics \& Intersubjectivity (CCI) Framework}.
  Internal research notes, 2026-07.
  \item Packer et al. \emph{MemGPT: Towards LLMs as Operating Systems.} 2023.
  \item Park et al. \emph{Generative Agents: Interactive Simulacra of Human Behavior.} 2023.
  \item Ying et al. \emph{GNNExplainer: Generating Explanations for Graph Neural Networks.} 2019.
  \item Shapiro et al. \emph{CRDTs: Consistency without consensus.} 2011.
\end{enumerate}

\paragraph*{Appendix --- Reproduction.}
Backend: \texttt{marvis\_liquid\_loop\_server.py} (8790).
Reference protocol impl: \texttt{liquid\_loop\_mesh\_v2.py} (\texttt{validate\_evidence},
\texttt{compute\_cci}, \texttt{cognitive\_health}, \texttt{detect\_conflict}). Spec:
\texttt{liquid\_loop\_mesh\_v2\_spec.md}. All code MIT, zero-dependency beyond stdlib +
liquid\_loop.

\end{document}
